How to improve your relationship with your future self, V 2.0

### 2016-06-28 workflow.Rmd
Version sent to Fernando and Thad

# Creating a reproducible workflow

@@

We all always collaborate. Many of us collaborate with groups of
people at one moment in time as we race against a deadline. All of us
collaborate with ourselves over time.\footnote{What is a reasonable
  time-span for which to plan for self-collaboration on a single idea?
  Ask your advisers how long it took them to move from idea to dissertation to
  publication.}  The time-frames over which collaboration are required
--- whether among a group of people working together or within a
single scholar's productive life or probably both --- are much longer
than any given version of any given software will easily exist. Plain
text is the exception. Thus, even as we extol version control systems,
one must have a way to ensure future access to them in a form that
will still be around when sentient cockroaches finally join political
science departments (by then dominated by cetaceans after humans are
mostly uploads).^[The arrival of the six-legged social
  scientists revives Emacs and finally makes Ctrl-a Ctrl-x Esc-x
  Ctrl-c a [reasonable key combination](http://kieran.healy.usesthis.com/).]

But what if no one ever hears of your work, or, by some cruel fate, your
article does not spawn debate? Why then would you spend time to
communicate with your future self and others? Our own answer to this
question is that we want our work to be credible and useful to ourself
and other scholars regardless.  What we report in our data analyses should have two main
characteristics: (1) the findings of the work should not be a matter
of opinion; and (2) other people should be able to reproduce the findings. That is,
the work represents a shared experience --- and an experience shared without respect to the
identities of others (although requiring some common technical
training and research resources).

Assume we want others to believe us when we say something. More narrowly,
assume we want other people to believe us when we say something about data:
'data' here can be words, numbers, musical notes, images, ideas, etc \ldots The
point is that we are making some claims about patterns in some collection of
stuff. For example, when Jake was invited into the homes and offices of ordinary
people in Chile in 1991, the "stuff" was recordings of conversations that we
had about life during the first year of democracy and Jake was comparing the
responses of people who had experienced the Pinochet dictatorship differently.
Now, it might be easy to convince others that 'this collection of stuff' is
different from 'that collection of stuff' if those people were looking over our
shoulders the whole time that we made decisions about collecting the stuff and
broke it up into understandable parts and reorganized and summarized it.
Unfortunately, we can't assume that people are willing to shadow a researcher
throughout her career. Rather, we do our work alone or in small groups and want
to convince other distant and future people about our analyses and findings.

Now, say your collections of stuff are large or complex and your
chosen tools of analyses are computer programs. How can we convince
people that what we did with some data with some program is credible,
not a matter of whim or opinion, and reproducible by others who didn't
shadow us as we wrote our papers? This essay has suggested a few
concrete ways to enhance the believability of such scholarly work. In
addition, these actions (as summarized in the section headings of this
essay) make collaboration within research groups more effective.
Believability comes in part from reproducibility and researchers
often need to be able to reproduce in part or in whole what different
people in the group have done or what they, themselves, did in the past.

In the end, following these practices and those recommended by
@fredrickson2011tpm and @healy2011tpm among others working on these topics
allows your computerized analyses of your collections of stuff to be
credible.  If then someone quibbles with your analyses, your future
self can shoot them the archive required to reproduce your
work.^[Since you used plain text, the files will still be
  intelligible, analyzed using commented code so that folks can
  translate to whatever system succeeds R, or since you used R, you
  can include a copy of R and all of the R packages you used in your
  final analyses in the archive itself. You can even throw in
  a copy of whatever version of linux you used and an open source virtual machine running the
  whole environment.] You can say, 'Here is everything you need to
reproduce my work.' To be extra helpful you can add 'Read the README
file for further instructions.' And then you can get on with your life:
maybe the next great idea will occur when your 4-year-old asks a wacky
question after stripping and painting her overly cooperative
1-year-old brother purple, or teaching a class, or in a coffee shop,
or on a quiet walk.

@@

<!-- dont forget Camerer 2016 Science and  Silberzahn and Uhlmann (Nature 2015)

>>>>

# Create a reproducible workflow

@@

We all always collaborate. Many of us collaborate with groups of people at one
moment in time as we race against a deadline. All of us collaborate with
ourselves over time.^[What is a reasonable time-span for which to plan
for self-collaboration on a single idea?  Ask your advisers how long it took
them to move from idea to dissertation to publication.]  The time-frames over
which collaboration are required --- whether among a group of people working
together or within a single scholar's productive life or probably both --- are
much longer than any given version of any given software will easily exist.
Plain text is the exception. Thus, even as we extol version control systems,
one must have a way to ensure future access to them in a form that will still
be around when sentient cockroaches finally join political science departments
(by then dominated by cetaceans after humans are mostly uploads).^[The arrival
of the six-legged social scientists revives Emacs and finally makes Ctrl-a
Ctrl-x Esc-x Ctrl-c a [reasonable key
combination](http://kieran.healy.usesthis.com/). So far, in terms of version
control systems, Jake has used RCS, CVS, bazaar, subversion, and now mostly
uses git.]

But what if no one ever hears of your work, or, by some cruel fate, your
article does not spawn debate? Why then would you spend time to communicate
with your future self and others? Our own answer to this question is that we
want our work to be credible and useful to ourselves and other scholars
regardless.  What we report in our data analyses should have two main
characteristics: (1) the findings of the work should not be a matter of
opinion; and (2) other people should be able to reproduce the findings. That
is, the work represents a shared experience --- and an experience shared
without respect to the identities of others (although requiring some common
technical training and research resources).

Assume we want others to believe us when we say something. More narrowly,
assume we want other people to believe us when we say something about data:
'data' here can be words, numbers, musical notes, images, ideas, etc. The point
is that we are making some claims about patterns in some collection of stuff.
For example, when Jake was invited into the homes and offices of ordinary
people in Chile in 1991, the "stuff" was recordings of conversations that he
had about life during the first year of democracy. Now, it might be easy to
convince others that 'this collection of stuff' is different from 'that
collection of stuff' if those people were looking over our shoulders the whole
time that we made decisions about collecting the stuff and broke it up into
understandable parts and reorganized and summarized it.  Unfortunately, we
can't assume that people are willing to shadow a researcher throughout her
career. Rather, we do our work alone or in small groups and want to convince
other distant and future people to take our analyses and findings seriously.

Now, say your collections of stuff are large or complex and your
chosen tools of analyses are computer programs. How can we convince
people that what we did with some data with some code is credible,
not a matter of whim or opinion, and reproducible by others who didn't
shadow us as we wrote our papers? This essay has suggested a few
concrete ways to enhance the believability of such scholarly work. In
addition, these actions (as summarized in the section headings of this
essay) make collaboration within research groups more effective.
Believability comes in part from reproducibility and researchers
often need to be able to reproduce in part or in whole what different
people in the group have done or what they, themselves, did in the past.

In the end, following these practices and those recommended by
@fredrickson2011tpm and @healy2011tpm among others working on these topics
allows your computerized analyses of your collections of stuff to be credible.
If then someone quibbles with your analyses, your future self can shoot them
the archive required to reproduce your work.^[Since you used plain text, the
files will still be intelligible, analyzed using commented code so that folks
can translate to whatever system succeeds R, or since you used R, you can
include a copy of R and all of the R packages you used in your final analyses
in the archive itself. You can even throw in a copy of whatever version of
linux you used and an open source virtual machine running the whole environment
using say, docker.] You can say, 'Here is everything you need to reproduce my
work.' To be extra helpful you can add 'Read the README file for further
instructions.' And then you can get on with your life: maybe the next great
idea will occur when your 4-year-old asks a wacky question after stripping and
painting her overly cooperative 1-year-old brother purple, or teaching a class,
or in a coffee shop, or on a quiet walk.

@@
